Describing the scale and composition of calls for police service: a replication and extension using open data

### 2022-02-11 paper.Rmd
update discussion
few changes + seperate feekback
feedback

Here, we propose an extension of @ratcliffe2021policing and @lum2021can. This extension has two principal components. First, we replicate the main analyses of these existing investigations into the scale and composition of police demand using open study materials. To date, investigations have been conducted on data which are not publicly available and without the materials (e.g., code) to reproduce the results. Open science practices have been identified as fundamental to the aims of evidence-based policing and vital in ensuring that in the long-term, real-life decisions are based on a concrete and replicable pool of evidence [@openedp2019]. In this case, the necessity for a fully open investigation into the scale and composition of police demand is particularly important given the public nature of the debate around police reform and the considerable implications of redefining the role of the police without transparent empirical investigation. It is reasonable for police practitioners and the public to expect a thorough, open and replicable investigation into police demand before academia concludes that the issue is resolved – certainly before findings are used to change or introduce police practice or public policy. Open materials have the additional benefit of being available for re-use in other jurisdictions, helping build a reliable and comparable evidence-base around the scale and composition of police demand across different contexts.

@@

The temporal patterning of the six demand type classifications are visualized using a heatmap of mean call counts, aggregated by day of the week, and hour of the day. To summarize the spatial patterning, call location coordinates are aggregated to synthetic grid cells using spatial joins. Grid cells are defined as 1000 ft$^2$ (305m$^2$) -- a size deemed appropriate based on a balance between capturing localized variation and interpretable visualizations. 2% of incidents are excluded from the spatial visualizations due to incomplete geographic information. To accurately convey the resource deployed to each area during the year, maps visualize the aggregate _time on scene_ (in hours) within each grid cell over the course of the year, for each demand type classification. The equivalent raw count maps are available in the Appendix. In this way, we visualize the spatio-temporal patterning of demand in a manner comparable to @ratcliffe2021policing, but here, we provide a comparison across different demand classifications and two different measures of demand.

@@

In this study, calls were categorized according the six classes used by @ratcliffe2021policing. The overall composition of police demand in terms of call volume and proportional time spent are remarkably similar. A considerable volume of calls for service, and in turn, a considerable proportion of deployed police time, relate to incident types which do not fall under the traditional 'crime-fighting' role of the police. Instead, the police fulfill a variety of other social service functions, including the resolution of public (mental) health incidents and resolving community issues. That said, descriptive statistics which drill down into the underlying distribution of officer time consumed by each demand type provide additional insight which would otherwise be overlooked. First, it is clear that there is considerable variation in the amount of time consumed both between and within demand classifications. On a methodological point, this highlights how average (e.g., mean times) or aggregate (e.g., sum times) might mask underlying variation. Second, and relatedly, these extensions demonstrate that a disproportionately large amount of police time is spent on a small number of emergency calls. 50% of total time on scene in consumed by just 15% of calls. High-demand calls will likely be complex and/or life-threatening incidents which require numerous officers to be present at the scene, or for extended periods of time. A targeted examination of these calls might shed light on the extent to which such high-intensity demands could be reduced or better resolved through or in concert with other public services. Targeted interventions on just a handful of calls could drastically reduce the public demand for police services.

>>>>

Here, we propose an extension of @ratcliffe2021policing and @lum2021can. This extension has two principal components. First, we replicate the main analyses of these existing investigations into the scale and composition of police demand using open study materials. To date, investigations have been conducted on data which are not publicly available and without the materials (e.g., code) to reproduce the results. Open science practices have been identified as fundamental to the aims of evidence-based policing and vital in ensuring that in the long-term, real-life decisions are based on a concrete and replicable pool of evidence [@openedp2019]. In this case, the necessity for a fully open investigation into the scale and composition of police demand is particularly important given the public nature of the debate around police reform and the considerable implications of redefining the role of the police without transparent empirical investigation. It is reasonable for police practitioners and the public to expect a thorough, open and replicable investigation into police demand before academia concludes that the issue is resolved – certainly before findings are used to change or introduce police practice or public policy. Especially since classifying police demand into these different categories is no straightforward or facile task. The difference in categories in Ratcliffe's and Lum's paper attest to this and showcase why they should not have to be taken at face-value. Open materials have the additional benefit of being available for re-use in other jurisdictions, helping build a reliable and comparable evidence-base around the scale and composition of police demand across different contexts.

@@

The temporal patterning of the six demand type classifications are visualized using a heatmap of mean call counts, aggregated by day of the week, and hour of the day. To summarize the spatial patterning, call location coordinates are aggregated to synthetic grid cells using spatial joins. Grid cells are defined as 1000 ft$^2$ (305m$^2$) -- a size deemed appropriate based on a balance between capturing localized variation and interpretable visualizations. 2% of incidents are excluded from the spatial visualizations due to incomplete geographic information. To accurately convey the resource deployed to each area during the year, maps visualize the aggregate _time on scene_ (in hours) within each grid cell over the course of the year, for each demand type classification. The equivalent raw count maps are available in the Appendix. In this way, we visualize the spatio-temporal patterning of demand in a manner comparable to @ratcliffe2021policing, while adding a comparison across different demand classifications and two different measures of demand.

@@

In this study, calls were categorized according the six classes used by @ratcliffe2021policing. The overall composition of police demand in terms of call volume and proportional time spent are remarkably similar. A considerable volume of calls for service, and in turn, a considerable proportion of deployed police time, relate to incident types which do not fall under the traditional 'crime-fighting' role of the police. Instead, the police fulfill a variety of other social service functions, including the resolution of public (mental) health incidents and resolving community issues. That said, descriptive statistics which drill down into the underlying distribution of officer time consumed by each demand type provide additional insight which would otherwise be overlooked. First, it is clear that there is considerable variation in the amount of time consumed both between and within demand classifications. On a methodological point, this highlights how average (e.g., mean times) or aggregate (e.g., sum times) might mask underlying variation. Second, and relatedly, these extensions demonstrate that a disproportionately large amount of police time is spent on a small number of emergency calls. 50% of total time on scene in consumed by just 15% of calls. High-demand calls will likely be complex and/or life-threatening incidents which require numerous officers to be present at the scene, or for extended periods of time. A targeted examination of these calls might shed light on the extent to which such high-intensity demands could be reduced or better resolved through or in concert with other public services.
